---
title: "Assignment"
output: html_document
date: "2024-11-23"
---

## QUESTION 01: Data Visualisation for Science Communication

*Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data. **Do not make a boxplot**.*

*Use the following references to guide you:*

-   [*https://www.nature.com/articles/533452a*](https://www.nature.com/articles/533452a){.uri}
-   [*https://elifesciences.org/articles/16800*](https://elifesciences.org/articles/16800){.uri}

*Note: Focus on visual elements rather than writing misleading text on it.*

### a) Provide your figure here:

```{r bad figure code, echo=FALSE}

library(here)
library(ggplot2)
library(palmerpenguins)
library(tidyverse)
write_csv(penguins_raw, here("data", "penguins_raw.csv"))
ggplot(penguins_raw, aes(x=`Body Mass (g)`, y=`Flipper Length (mm)`)) + geom_smooth(method = "lm", se = FALSE, colour="lightgrey") +
  theme(axis.title = element_text(size = 5), axis.text = element_text(size = 5))

# There is no need to provide the code for your bad figure, just use echo=FALSE so the code is hidden. Make sure your figure is visible after you knit it. 

```

### b) Write about how your design choices mislead the reader about the underlying data (100-300 words).

# This figure is a linear regression of flipper length against body mass. The first misleading design choice is that the data used for the linear regression comes from three different species of penguins. Even though the regression is accurate, the lack of differentiation of species may hide any underlying species-specific trends, hence misleading readers into thinking that the correlation between body mass and flipper length can be generalised between species even though this is not justified. A better design choice may be to conduct three separate linear regressions, one for each species, and colour them differently, with a legend to label each regression line. Additionally, variation in the data is not shown in the figure. The raw data points were not plotted in the figure and the confidence interval of the regression line is not plotted. These two poor design choices leave the readers without information on the variation in the data. Plotting the raw data uncovers the underlying data and makes it accessible to the readers. Furthermore, the text size of the axes titles and numbers are extremely small and difficult to read. This may cause mistakes in reading the variables and the values in the regression. Increasing the text sizes would imrpove the figure. Lastly, the regression line is plotted in grey against a grey background. This design choice may result in mistakes in identifying the linear regression. A better design choice would choose a more contrasting colour for the regression line so that it stands out against the background.

------------------------------------------------------------------------

## QUESTION 2: Data Pipeline

*Write a data analysis pipeline in your .rmd RMarkdown file. You should be aiming to write a clear explanation of the steps as well as clear code.*

*Your code should include the steps practiced in the lab session:*

-   *Load the data*

-   *Appropriately clean the data*

-   *Create an Exploratory Figure (**not a boxplot**)*

-   *Save the figure*

-   ***New**: Run a statistical test*

-   ***New**: Create a Results Figure*

-   *Save the figure*

*An exploratory figure shows raw data, such as the distribution of the data. A results figure demonstrates the stats method chosen, and includes the results of the stats test or model.*

*Between your code, communicate clearly what you are doing and why.*

*Your text should include:*

-   *Introduction*

-   *Hypothesis*

-   *Stats Method*

-   *Results*

-   *Discussion*

-   *Conclusion*

*You will be marked on the following:*

### a) Your code for readability and functionality

### b) Your figures for communication

### c) Your text communication of your analysis

*Below is a template you can use.*

------------------------------------------------------------------------

### Introduction

# Data pipelines are important in ensuring science and figures are reproducible. Many current research papers present their experimental methodology, but not their methodology in data analysis and figures. This section of the assignment presents a data anlysis pipeline using the Palmer Penguin dataset as an example. It details the precise steps taken when converting raw data into clean data, analysing the cleaned data, and producing figures.

# I start by loading and cleaning the data

```{r Data Cleaning}

#load relevant packages
library(tidyverse)
library(palmerpenguins) #dataset comes from this package
library(here)
library(janitor)
library(ggplot2)
library(svglite)

source(here("functions", "cleaning.R")) #load functions that clean data

#load the data
write_csv(penguins_raw, here("data", "penguins_raw.csv")) #csv file written in the folder "data"
penguins_raw <- read_csv(here("data", "penguins_raw.csv"), show_col_types = FALSE) #loads the raw data

#clean data
penguins_clean <- penguins_raw %>% 
  clean_column_names() %>%          #clean column names
  remove_columns("comments") %>%    #remove comments column
  remove_columns("delta") %>%       #remove delta columns
  remove_empty_columns_rows %>%     #remove empty columns and rows
  shorten_species()                 #shorten species names

#save and load the clean data
write_csv(penguins_clean, here("data", "penguins_clean.csv")) #saves cleaned data in the "data" folder
penguins_clean <- read_csv(here("data", "penguins_clean.csv"), , show_col_types = FALSE) #load the cleaned data

```

# Having looked through the cleaned penguin data, I am interested in exploring the potential relationship between body mass and flipper length in the three species of penguins. Hence, I will produce an exploratory figure that examines these two variables and save the figure.

```{r Data Exploration}
#subset the data
subset <- penguins_clean %>% 
  select(species, flipper_length_mm, body_mass_g, )  %>% #subset out species, flipper length, and body mass
  remove_NA()                                            #remove rows with missing values

#define species colours
species_colours <- c("Adelie" = "darkorange",
                     "Chinstrap" = "purple",
                     "Gentoo" = "cyan4")

#produce exploratory figure
exploratory_figure <- ggplot(subset, aes(x = body_mass_g, y = flipper_length_mm, colour = species)) +
  geom_point() + #produce scatterplot
  scale_color_manual(values = species_colours) + #change species colours
  labs(
    x = "Body mass (g)", #x-axis label
    y = "Flipper length (mm)", #y-axis label
    colour = "Species"
  ) +
  theme_bw() #clean aesthetics
exploratory_figure

#save exploratory figure
svglite("figures/exploratory_figure.svg",
        width = 15,
        height = 10,
        scaling = 2)
print(exploratory_figure)
dev.off()

```

### Hypothesis

#From the exploratory figure above, we observe that flipper length and body mass are positively correlated in all three species of penguins. The Adelie penguins seem to have more dispersed data points and hence a larger variation in the data. Hence, it would be interesting to test if there is a significant linear relationship between body mass and flipper length in the Adelie penguins.
#Null hypothesis: there is no significant relationship between body mass and flipper length in Adelie penguins
#Alternative hypothesis: there is a significant relationship between body mass and flipper length in Adelie penguins

### Statistical Methods

#To test if there is a significant relationship between body mass and flipper length in Adelie penguins, I start by filtering the Adelie penguin data from the cleaned dataset. Following which, I fit a linear regression model to predict flipper length using body mass. Finally, I produce diagnostic plots to test if the assumptions of a linear regression are met.

```{r Statistics}

#filter Adelie penguin data
adelie_data <- penguins_clean %>% 
  select(species, flipper_length_mm, body_mass_g, )  %>% #subset out species, flipper length, and body mass
  remove_NA() %>%                                       #remove rows with missing values
  filter(species == "Adelie")                           #filter Adelie penguin data only

#statistical analysis
adelie_regression <- lm(flipper_length_mm ~ body_mass_g, data = adelie_data) #linear regression model
summary(adelie_regression) #output of linear model

#check assumptions
par(mfrow = c(2, 2)) #create a 2 by 2 grid for diagnostic plots
plot(adelie_regression) #produce diagnostic plots

```
### Results & Discussion

#Having run the statistical analysis, the following code is used to produce a results figure, which will be discussed below.

```{r Plotting Results}

#save adjusted R-squared value
r_squared <- summary(adelie_regression)$adj.r.squared

#produce results figure
results_figure <- ggplot(adelie_data, aes(x=body_mass_g, y=flipper_length_mm)) +
  geom_point(colour="darkorange") +                       #scatter plot
  geom_smooth(method = "lm", se = TRUE, color="red3") +   #linear regression
  labs(
    x = "Body mass (g)",       #x-axis label
    y = "Flipper length (mm)"  #y-axis label
  ) +
  annotate("text", x = 4600, y = 212, label = paste("R² =", round(r_squared, 3)), size = 5, color = "red3") +                      #add R-squared value
  theme_bw()                   #clean aesthetics
results_figure

#save results figure
svglite("figures/results_figure.svg",
        width = 15,
        height = 10,
        scaling = 2)
print(results_figure)
dev.off()

```

#From the output of the linear model, there is a significant linear relationship between body mass and flipper length in Adelie penguins (t=6.47, df=149, p=1.34e-09). Assumptions of a linear regression model are met based on the diagnostics plots. They show that the residuals follow a normal distribution, have equal variance, and there are no influential cases (outliers). Hence, we can reject the null hypothesis and conclude that there is a significant relationship between body mass and flipper length in Adelie penguins.

#However, the adjusted R-squared value of 0.214 is considerably low, meaning that only about 21% of the variation in flipper length can be explained by the variation in body mass. Hence, there may be other more important variables affecting flipper length, such as height. Furthermore, there could be sex differences that were not addressed by this analysis as the dataset contained measurements from both male and female Adelie penguins.

### Conclusion

#This analysis explored the correlation between body mass and flipper length in penguins using the Palmer Penguin dataset. From the exploratory figure, I decided to analyse the linear relationship between these two variables in the Adelie penguins. Using a linear regression model, I showed that there is a significant linear relationship between body mass and flipper length in Adelie penguins. However, the low adjusted R-squared value showed that the variation in flipper length cannot be fully explained by the variation in body mass. Hence, further investigation into other factors that determine flipper length in Adelie penguins is needed.

------------------------------------------------------------------------

## QUESTION 3: Open Science

### a) GitHub

*Upload your RProject you created for **Question 2** and any files and subfolders used to GitHub. Do not include any identifiers such as your name. Make sure your GitHub repo is public.*

*GitHub link:* https://github.com/benaiahhoong/ReproducibleAssignment/tree/main

*You will be marked on your repo organisation and readability.*

### b) Share your repo with a partner, download, and try to run their data pipeline.

*Partner's GitHub link:*

*You **must** provide this so I can verify there is no plagiarism between you and your partner.*

### c) Reflect on your experience running their code. (300-500 words)

-   *What elements of your partner's code helped you to understand and run their data pipeline?*

-   *Did it run? Did you need to fix anything?*

-   *What suggestions would you make for improving their code to make it more understandable or reproducible, and why?*

-   *If you needed to alter your partner's figure using their code, do you think that would be easy or difficult, and why?*

### d) Reflect on your own code based on your experience with your partner's code and their review of yours. (300-500 words)

-   *What improvements did they suggest, and do you agree?*

-   *What did you learn about writing code for other people?*

### e) What are the main barriers for scientists to share their data and code, and what could be done to overcome them? (500-700 words)

- Maitner et al. Code sharing increases citations, but remains uncommon. https://doi.org/10.21203/rs.3.rs-3222221/v1
- Trisovic et al. A large-scale study on research code quality and execution. https://rdcu.be/dZB7x
- A Rock-Star Researcher Spun a Web of Lies—and Nearly Got Away with It. https://thewalrus.ca/a-rock-star-researcher-spun-a-web-of-lies-and-nearly-got-away-with-it/
- Gomes et al. Why don't we share data and code? Perceived barriers and benefits to public archiving practices https://doi.org/10.1098/rspb.2022.1113